Demystifying Double Robust, Flexible Adjustment Methods for Causal Inference

### 2023-02-20 draft/ms-adj-draft.Rmd
added lalonde

```{r load}
# data from mx0.5
pisa <- readRDS('/Users/nathan/Google Drive/1 UCLA/Dissertation/diss-prop/data/pisa.rds') %>%
  mutate(age = as.numeric(age))

@@

# Preliminary Findings

## Monte Carlo Simulations

The structure of these simulations is based on @kang_2007_demystifying. For each unit $i = 1, \ldots, n$, let $Z = (z_{i1}, z_{i2}, z_{i3}, z_{i4})^\top$ be distributed independently as $N(0,I)$, where $I$ is the $4 \times 4$ identity matrix. Furthermore, let $d_i \in \{0,1\}$ be an indicator for treatment status with a Bernoulli distribution with probability $\pi_i$ of receiving treatment status. Values of $\pi_i$ are generated as:
$$\pi_i = \text{expit}(z_{i1} +0.5z_{i2}-0.33z_{i3} -0.2z_{i4}),$$
where $\text{expit}(\cdot) = \exp(\cdot)/(1 + \exp(\cdot)).$ These are then used as the probability of treatment assignment in a series of Bernoulli draws for $d_i$. Outcomes $y_i$ are generated as
$$y_i = 500 + 50d_i + 30z_{i1} -35z_{i2} - 60z_{i3} + 50z_{i4} + \epsilon_i,$$
where $\epsilon_i \sim N(0,10)$. Now assume that the researcher cannot measure the $z_{ij}$'s. Instead we observe $X = (x_{i1}, x_{i2}, x_{i3}, x_{i4})^\top$ :
$$\begin{aligned}
x_{i1} &= \exp(z_{i1}/2) \\
x_{i2} &= z_2 / (1+\exp(z_{i1})) + 10 \\
x_{i3} &= (z_{i1} z_{i3} /25 + 0.6)^3 \\
x_{i4} &= (z_2 + z_4 + 20)^2 \\
\end{aligned}$$

For these preliminary results, datasets of 200 observations are generated 100 times. For each estimation method, three specifications are used:

- "correct": the model is fitted to the true data-generating variables, $Z$
- "incorrect": the model is fitted to the transformed variables $X$
- "ovb": the model is fitted to $(x_2, x_3, x_4)^\top$ ($x_1$ is omitted).

Results are shown in Table \@ref(tab:sim-results). Estimation methods include ordinary least squares regression ("OLS"),  propensity-score matching with scores estimated from logistic regression using the `MatchIt` package ("PSM"), the augmented inverse propensity weighted estimator with generalized additive models using the `CausalGAM` package (AIPW), targeted maximum likelihood estimation using the `tmle` package (TMLE), and double/debiased machine learning with random forests with the `DoubleML` and `mlr3` packages.

Surprisingly, the double robust methods do not necessarily perform better. All methods perform well when the true data-generating variables are provided. When transformed variables are provided instead, bias, RMSE, and MAE rise for all of the estimation methods, even those that use flexible methods to model the response and treatment assignment surfaces. DML performs relatively well but RMSE and MAE still rise somewhat. When a variable is omitted, all of the methods perform poorly.

```{r sim}
expit <- function(x){exp(x) / (1+exp(x))}

sim_n <- function(n, mu = c(0, 0, 0, 0), Sigma = diag(c(1,1,1,1))){
  z <- MASS::mvrnorm(n, mu, Sigma)
  z1 <- z[,1]
  z2 <- z[,2]
  z3 <- z[,3]
  z4 <- z[,4]
  epsilon <- rnorm(n, 0, 10)

  pi <- expit(z1 + 0.5*z2 -.33*z3 - .2*z4)
  d <- rbinom(n, 1, pi)

  y <- 475 + 50*d + 30*z1 -35*z2 - 60*z3 + 50*z4 + epsilon

  x1 <- exp(z1/2)
  x2 <- z2 / (1+exp(z1)) + 10
  x3 <- (z1 * z3 / 25 + 0.6)^3
  x4 <- (z2 * z4 + 20)^2
  return(data.frame(
    z1 = z1,
    z2 = z2,
    z3 = z3,
    z4 = z4,
    pi = pi,
    d = d,
    y = y,
    x1 = x1,
    x2 = x2, 
    x3 = x3,
    x4 = x4
  ))
}

set.seed(1858)
sim_200 <- replicate(100, sim_n(200), simplify = F)
sim_1000 <- replicate(100, sim_n(1000), simplify = F)
sim_10000 <- replicate(100, sim_n(10000), simplify = F)
```

```{r lm-sim, eval = F}
lm_sim <- function(dat){
  n <- length(dat)
  lm_list_true <- list()
  lm_list_false <- list()
  lm_list_ovb <- list()
  for(i in 1:n){
    lm_out_true <- lm(y ~ d + z1 + z2 + z3 + z4, dat[[i]])
    lm_list_true[[i]] <- lm_out_true$coefficients[['d']]

    lm_out_false <- lm(y ~ d + x1 + x2 + x3 + x4, dat[[i]])
    lm_list_false[[i]] <- lm_out_false$coefficients[['d']]

    lm_out_ovb <- lm(y ~ d + x2 + x3 + x4, dat[[i]])
    lm_list_ovb[[i]] <- lm_out_ovb$coefficients[['d']]
  }

  return(list(
    n = nrow(dat[[i]]),
    correct = unlist(lm_list_true),
    incorrect = unlist(lm_list_false),
    ovb = unlist(lm_list_ovb)))
}

perform <- function(est_list, label){
  perform_list <- list()

  for(i in 1:(length(est_list)-1)){
    est <- unlist(est_list[[i+1]])
    bias <- mean(est-50)
    percent_bias <- 100 * bias / sd(est)
    rmse <- sqrt(mean((est - 50)^2))
    mae <- median(abs(est - 50))

    perform_list[[i]] <- data.frame(label = label, n = est_list[['n']], 
                                    model = names(est_list)[i+1], 
                                    bias, percent_bias, rmse, mae)
  }
  return(bind_rows(perform_list))
}

lm_sim_200 <- lm_sim(sim_200) 
# lm_sim_1000 <- lm_sim(sim_1000) 
# lm_sim_10000 <- lm_sim(sim_10000)

lm_df <- bind_rows(
  perform(lm_sim_200, 'OLS'),
  )

write_csv(lm_df, here('files', 'lm_df.csv'))
```

```{r psm-sim, eval = F}
psm_sim <- function(dat){
  n <- length(dat)
  psm_list_true <- list()
  psm_list_false <- list()
  lm_list_ovb <- list()
  for(i in 1:n){
    match_true <- MatchIt::matchit(d ~ z1 + z2 + z3 + z4,
                           data = dat[[i]],
                           method = 'nearest',
                           distance = 'glm')

    psm_out_true <- lm(y ~ d + z1 + z2 + z3 + z4, 
                       MatchIt::match.data(match_true), 
                       weights = weights)

    match_false <- MatchIt::matchit(d ~ x1 + x2 + x3 + x4,
                           data = dat[[i]],
                           method = 'nearest',
                           distance = 'glm')

    psm_out_false <- lm(y ~ d + x1 + x2 + x3 + x4, 
                       MatchIt::match.data(match_false), 
                       weights = weights)

    match_ovb <- MatchIt::matchit(d ~ x2 + x3 + x4,
                           data = dat[[i]],
                           method = 'nearest',
                           distance = 'glm')

    psm_out_ovb <- lm(y ~ d + x2 + x3 + x4, 
                       MatchIt::match.data(match_false), 
                       weights = weights)

    psm_list_true[[i]] <- psm_out_true$coefficients[['d']]
    psm_list_false[[i]] <- psm_out_false$coefficients[['d']]
    psm_list_ovb[[i]] <- psm_out_ovb$coefficients[['d']]
  }

  return(list(n = nrow(dat[[i]]),
              correct = unlist(psm_list_true),
              incorrect = unlist(psm_list_false),
              ovb = unlist(psm_list_ovb)))
}

psm_sim_200 <- psm_sim(sim_200) 
# psm_sim_1000 <- psm_sim(sim_1000) 
# psm_sim_10000 <- psm_sim(sim_10000)

psm_df <- bind_rows(
  perform(psm_sim_200, 'PSM'),
  )

write_csv(psm_df, here('files', 'psm_df.csv'))
```

```{r aipw-sim, eval = F}
aipw_sim <- function(dat, seed = 185){
  library(gam)

  set.seed(seed)
  n <- length(dat)
  aipw_list_true <- list()
  aipw_list_false <- list()
  aipw_list_ovb <- list()

  for(i in 1:n){

    print(i)

    correct <- CausalGAM::estimate.ATE(
        pscore.formula = d ~ s(z1) + s(z2) + s(z3) + s(z4),
        pscore.family = binomial,
        outcome.formula.t = y ~ s(z1) + s(z2) + s(z3) + s(z4),
        outcome.formula.c = y ~ s(z1) + s(z2) + s(z3) + s(z4),
        outcome.family = gaussian,
        treatment.var = 'd',
        data=dat[[i]],
        divby0.action="t",
        divby0.tol=0.001,
        var.gam.plot=FALSE,
        nboot=50)

    incorrect <- CausalGAM::estimate.ATE(
        pscore.formula = d ~ s(x1) + s(x2) + s(x3) + s(x4),
        pscore.family = binomial,
        outcome.formula.t = y ~ s(x1) + s(x2) + s(x3) + s(x4),
        outcome.formula.c = y ~ s(x1) + s(x2) + s(x3) + s(x4),
        outcome.family = gaussian,
        treatment.var = 'd',
        data=dat[[i]],
        divby0.action="t",
        divby0.tol=0.001,
        var.gam.plot=FALSE,
        nboot=50)

    ovb <- CausalGAM::estimate.ATE(
        pscore.formula = d ~ s(x2) + s(x3) + s(x4),
        pscore.family = binomial,
        outcome.formula.t = y ~ s(x2) + s(x3) + s(x4),
        outcome.formula.c = y ~ s(x2) + s(x3) + s(x4),
        outcome.family = gaussian,
        treatment.var = 'd',
        data=dat[[i]],
        divby0.action="t",
        divby0.tol=0.001,
        var.gam.plot=FALSE,
        nboot=50)

    aipw_list_true[[i]] <- correct$ATE.AIPW.hat
    aipw_list_false[[i]] <- incorrect$ATE.AIPW.hat
    aipw_list_ovb[[i]] <- ovb$ATE.AIPW.hat
  }

  return(list(n = nrow(dat[[i]]),
              correct = unlist(aipw_list_true),
              incorrect = unlist(aipw_list_false),
              ovb = unlist(aipw_list_ovb)))
}

aipw_out <- aipw_sim(sim_200)

aipw_df <- perform(aipw_out, 'AIPW')

write_csv(aipw_df,  here('files', 'aipw_df.csv'))

@@

tmle_sim <- function(dat){
  n <- length(dat)

  tmle_list_true_true <- list()
  tmle_list_false_true <- list()
  tmle_list_true_false <- list()
  tmle_list_false_false <- list()
  tmle_list_ovb_ovb <- list()

  for(i in 1:n){

    print(i)

    tmle_false_false <- with(
      dat[[i]],
      tmle::tmle(
        Y = y,
        A = d,
        W = data.frame(z1, z2, z3, z4, x1, x2, x3, x4),
        Qform = Y ~ A + x1 + x2 + x3 + x4,
        gform = A ~ x1 + x2 + x3 + x4))

    tmle_false_true <- with(
      dat[[i]],
      tmle::tmle(
        Y = y,
        A = d,
        W = data.frame(z1, z2, z3, z4, x1, x2, x3, x4),
        Qform = Y ~ A + x1 + x2 + x3 + x4,
        gform = A ~ z1 + z2 + z3 + z4))

    tmle_true_false <- with(
      dat[[i]],
      tmle::tmle(
        Y = y,
        A = d,
        W = data.frame(z1, z2, z3, z4, x1, x2, x3, x4),
        Qform = Y ~ A + z1 + z2 + z3 + z4,
        gform = A ~ x1 + x2 + x3 + x4))

    tmle_true_true <- with(
      dat[[i]],
      tmle::tmle(
        Y = y,
        A = d,
        W = data.frame(z1, z2, z3, z4, x1, x2, x3, x4),
        Qform = Y ~ A + z1 + z2 + z3 + z4,
        gform = A ~ z1 + z2 + z3 + z4))

    tmle_ovb_ovb <- with(
      dat[[i]],
      tmle::tmle(
        Y = y,
        A = d,
        W = data.frame(x2, x3, x4),
        Qform = Y ~ A + x2 + x3 + x4,
        gform = A ~ x2 + x3 + x4))

    tmle_list_true_true[[i]] <- tmle_true_true$estimates$ATE$psi
    tmle_list_true_false[[i]] <- tmle_true_false$estimates$ATE$psi
    tmle_list_false_true[[i]] <- tmle_false_true$estimates$ATE$psi
    tmle_list_false_false[[i]] <- tmle_false_false$estimates$ATE$psi
    tmle_list_ovb_ovb[[i]] <- tmle_ovb_ovb$estimates$ATE$psi
  }

  return(list(n = nrow(dat[[i]]),
              correct_correct = unlist(tmle_list_true_true),
              correct_incorrect = unlist(tmle_list_true_false),
              incorrect_correct = unlist(tmle_list_false_true),
              incorrect_incorrect = unlist(tmle_list_false_false),
              ovb = unlist(tmle_list_ovb_ovb))
         )
}

tmle_out <- tmle_sim(sim_200)

tmle_df <- perform(tmle_out, 'TMLE')

write_csv(tmle_df, here('files', 'tmle_df.csv'))
```

@@

@@

  correct_list <- list()
  incorrect_list <- list()
  ovb_list <- list()

  for(i in 1:n){
    print(i)

@@

    correct = DoubleML::DoubleMLPLR$new(
      DoubleML::DoubleMLData$new(dat[[i]],
                                 y_col = 'y',
                                 d_cols = 'd',
                                 x_cols = c('z1', 'z2', 'z3', 'z4')), 
      ml_l=ml_l, ml_m=ml_m)
    correct$fit()
    correct_list[[i]] <- correct$all_coef[[1,1]]

    incorrect = DoubleML::DoubleMLPLR$new(
      DoubleML::DoubleMLData$new(dat[[i]],
                                 y_col = 'y',
                                 d_cols = 'd',
                                 x_cols = c('x1', 'x2', 'x3', 'x4')), 
      ml_l=ml_l, ml_m=ml_m)
    incorrect$fit()
    incorrect_list[[i]] <- incorrect$all_coef[[1,1]]

    ovb = DoubleML::DoubleMLPLR$new(
      DoubleML::DoubleMLData$new(dat[[i]],
                                 y_col = 'y',
                                 d_cols = 'd',
                                 x_cols = c('x2', 'x3', 'x4')), 
      ml_l=ml_l, ml_m=ml_m)
    ovb$fit()
    ovb_list[[i]] <- ovb$all_coef[[1,1]]

  }
  return(list(n = nrow(dat[[i]]),
              correct = unlist(correct_list),
              incorrect = unlist(incorrect_list),
              ovb = unlist(ovb_list))
         )
}

dml_out <- dml_sim(sim_200)

dml_df <- perform(dml_out, 'DML')

write_csv(dml_df, here('files', 'dml_df.csv'))
```

```{r sim-results}
bind_rows(
  read_csv(here('files', 'lm_df.csv')),
  read_csv(here('files', 'psm_df.csv')),
  read_csv(here('files', 'aipw_df.csv')),
  read_csv(here('files', 'tmle_df.csv')) %>%
    filter(model %in% c('correct_correct', 'incorrect_incorrect', 'ovb')) %>%
    mutate(model = c('correct', 'incorrect', 'ovb')),
  read_csv(here('files', 'dml_df.csv'))
  ) %>%
  kableExtra::kable(booktabs = T, 
                    linesep = '',
                    caption = 'Results of Monte Carlo simulations over 100 replications. Percent bias is calculated as the estimator\'s bias as a percentage of its standard error, rmse is root mean squared error, and mae is median absolute error.')
```

```{r test-sample, eval = F}
set.seed(185)
pisa_test <- bind_rows(
  pisa %>%
    filter(match_country == 'Albania', immigrant == T) %>%
    select(math1, immigrant, mom_ed, dad_ed,
                          female, early_ed, cultural_pos, home_ed, age) ,
    #sample_n(200),
  pisa %>%
    filter(match_country == 'Albania', immigrant == F) %>%
        select(math1, immigrant, mom_ed, dad_ed,
                          female, early_ed, cultural_pos, home_ed, age) 
    #sample_n(200)
  )

# pisa_test <- pisa %>%
#     filter(match_country == 'Albania') %>%
#  select(math1, immigrant, mom_ed, dad_ed,
#                          female, early_ed, cultural_pos, home_ed, age)

#  filter(match_country %in% c('Portugal', 'Germany', 'United States'))
```

```{r tmle, eval = F}
# Y = outcome
# A = binary treatment
# W = covariates

tmle_out <- with(pisa_test, tmle::tmle(
                Y = math1,
                A = as.logical(immigrant),
                W = select(pisa_test, mom_ed, dad_ed, 
                          female, early_ed, cultural_pos, home_ed, age),
                verbose = T))
tmle_out

# pisa %>%
#  count(match_country, immigrant) %>% View()
#   arrange(desc(n))

```{r dml, eval = F}
dml_data <- DoubleML::DoubleMLData$new(as.data.frame(pisa_test),
                y_col = 'math1',
                d_cols = 'immigrant',
                x_cols = c('mom_ed', 'dad_ed', 'female', 'early_ed', 
                'cultural_pos', 'home_ed', 'age'))

library(mlr3)
library(mlr3learners)
# surpress messages from mlr3 package during fitting
lgr::get_logger("mlr3")$set_threshold("warn")

learner = lrn("regr.ranger", num.trees=500, 
              max.depth=5, min.node.size=2)
ml_g = learner$clone()
ml_m = learner$clone()

set.seed(185)

obj_dml_plr = DoubleML::DoubleMLPLR$new(dml_data, ml_g=ml_g, ml_m=ml_m)
obj_dml_plr$fit()
print(obj_dml_plr)
```

```{r plce, eval = F}
plce_out <- with(pisa_test, PLCE::plce(
                y = math1,
                treat = immigrant,
                X = as.matrix(select(pisa_test, mom_ed, dad_ed, #early_ed,
                          female, cultural_pos, home_ed, age))))

```{r bart, eval = F}
set.seed(159)
bart_out <- bartCause::bartc(
  response = pisa_test$math1,
  treatment = pisa_test$immigrant,
  confounders = as.data.frame(select(pisa_test, mom_ed, dad_ed,
                          female, early_ed, cultural_pos, home_ed, age))
  )

summary(bart_out)

>>>>

```{r, functions}
perform <- function(est_df, label){
  est_df %>%
    summarize(bias = mean(d - truth),
              percent_bias = bias/sd(d),
              rmse = sqrt(mean(d - truth)^2),
              mae = median(abs(d - truth)),
              fail_count = first(fail_count),
              comp_time = first(comp_time)) %>%
    mutate(label = label,
           n = nrow(est_df) + fail_count) %>%
    select(label, everything()) %>%
    return()
}
```

```{r load, eval = F}
# data from mx0.5
pisa <- readRDS('/Users/nathan/My Drive/1 UCLA/Dissertation/data/pisa.rds') %>%
  mutate(age = as.numeric(age))

@@

# Methods

For the comparisons in the next section, estimation methods include ordinary least squares regression ("OLS"),  propensity-score matching with scores estimated from logistic regression using the `MatchIt` package ("PSM"), the augmented inverse propensity weighted estimator with generalized additive models using the `CausalGAM` package (AIPW), targeted maximum likelihood estimation using the `tmle` package (TMLE) with the default Superlearner that relies on the `SL.glm`,` SL.step`, and `GL.glm.interaction` functions, and double/debiased machine learning with random forests with the `DoubleML` and `mlr3` packages.

# Results

## Simluation with Dorie et al. (2019) Data

In 2016, the Atlantic Causal Inference Conference hosted a competition for causal inference methods that adjust on observables. @dorie_2019_automated published the results of this competition, along with the data used in the competition. Below, I test double robust methods on the 20 data sets used for the "do-it-yourself" part of the competition. The data represent a hypothetical twins study investigating the impact of birth weight on IQ. The data have 4802 observations and 52 covariates. The authors of the study specify a different data generating process for the potential outcomes in each data set. In all cases, ignorability holds, but the authors vary the following:

- degree of nonlinearity
- percentage of treated
- overlap for the treatment group
- alignment (correspondence between the assignment mechanism and the response surface)
- treatment effect heterogeneity
- overall magnitude of the treatment effect

The 20 data sets used here cover a range of these attributes; see the supplemental material from @dorie_2019_automated for details. For these preliminary results, I used 5 simulations of each data set, resulting in 100 data sets. I then calculated bias, percent bias (the estimator's bias as a percentage of its standard error), root mean squared error (rmse), and median absolute error (mae). I also include the number of data sets for which the method failed and the average computation time for each data set, in seconds.

```{r dorie-sim, eval = F}
set.seed(185)
sim_list <- list()
for(i in 1:20){ # up to 77
  #sim_list_nest <- list()
  for(j in 1:10){ # up to 100
    print(paste0(i,',',j))
    sim_list[[paste0(i,',',j)]] <- aciccomp2016::dgp_2016(aciccomp2016::input_2016, 
                                      i, j, extraInfo = T)
  }
  #sim_list[[i]] <- sim_list_nest
}

# dat <- sim_list
# i <- '3,1'
# 
# sim_dat <- data.frame(y = dat[[i]]$y, 
#                           d = ifelse(dat[[i]]$z == 'trt', 1, 0), 
#                           dat[[i]]$x)
# lm_out <- lm(y ~ d + ., sim_dat)
# lm_list[[i]] <- data.frame(d = lm_out$coefficients[['d']],
#                            truth = mean(dat[[i]]$y.1) - mean(dat[[i]]$y.0))

sim_linear <- list()
for(i in c(1, 3))
  for(j in 1:100){ 
    print(paste0(i,',',j))
    sim_linear[[paste0(i,',',j)]] <- aciccomp2016::dgp_2016(aciccomp2016::input_2016, 
                                      3, j, extraInfo = T)
  }
}

```{r lm-sim, eval = F}
lm_sim <- function(dat){
  start_time <- Sys.time()
  n <- length(dat)
  lm_list <- list()
  for(i in 1:n){
    print(paste0(i, ' out of ', n))
    sim_dat <- data.frame(y = dat[[i]]$y, 
                          d = ifelse(dat[[i]]$z == 'trt', 1, 0), 
                          dat[[i]]$x)

    tryCatch({
      lm_out <- tidy(lm(y ~ d + ., sim_dat))
      lm_list[[i]] <- data.frame(d = lm_out[[2,2]],
                                 se = lm_out[[2,3]],
                                 truth = mean(dat[[i]]$y.1) - mean(dat[[i]]$y.0))
      }, error=function(e){
        cat("ERROR :",conditionMessage(e), "\n")
        })

  }

  fail_count <- sum(sapply(lm_list, function(x) is.null(x)))
  # in case last few are errors
  fail_count <- ifelse(length(lm_list) == n, fail_count, fail_count + (n - length(lm_list)))

  end_time <- Sys.time()

  return(list(
    est_df = bind_rows(lm_list),
    fail_count = fail_count,
    comp_time = as.numeric(difftime(end_time, start_time, units = 'secs'))/(n-fail_count))
    )
}

lm_sim_out <- lm_sim(sim_list) 
lm_sim_linear <- lm_sim(sim_linear)

lm_df <- lm_sim_out$est_df %>%
  mutate(fail_count = lm_sim_out$fail_count,
         comp_time = lm_sim_out$comp_time)

lm_df_linear <- lm_sim_linear$est_df %>%
  mutate(fail_count = lm_sim_linear$fail_count,
         comp_time = lm_sim_linear$comp_time)

write_csv(lm_df, here('files', 'lm_df.csv'))
write_csv(lm_df_linear, here('files', 'lm_df_linear.csv'))

# lm_lalonde <- lm_sim(lalonde)

```{r psm-sim, eval = F}
psm_sim <- function(dat){
  start_time <- Sys.time()
  n <- length(dat)
  psm_list <- list()
  for(i in 1:n){
    print(paste0(i, ' out of ', n))
    sim_dat <- data.frame(y = dat[[i]]$y, 
                          d = ifelse(dat[[i]]$z == 'trt', 1, 0), 
                          dat[[i]]$x)

    form <- as.formula(paste0('d ~ ', paste(names(dat[[i]]$x), collapse = '+')))
    match_out <- MatchIt::matchit(form,
                           data = sim_dat,
                           method = 'nearest',
                           distance = 'glm')

    # match_data <- MatchIt::match.data(match_out) 
    # apply(match_data, 1, unique)
    form2 <- as.formula(paste0('y ~ d + ', paste(names(dat[[i]]$x), collapse = '+')))

  tryCatch({
    psm_out <- lm(form2, 
                  MatchIt::match.data(match_out), 
                  weights = weights) %>%
      tidy()
    psm_list[[i]] <- data.frame(d = psm_out[[2,2]],
                                se = psm_out[[2,3]],
                                truth = mean(dat[[i]]$y.1) - mean(dat[[i]]$y.0))

  }, error=function(e){
    cat("ERROR :",conditionMessage(e), "\n")
    })

  }

  fail_count <- sum(sapply(psm_list, function(x) is.null(x)))
  # in case last few are errors
  fail_count <- ifelse(length(psm_list) == n, fail_count, fail_count + (n - length(psm_list)))

  end_time <- Sys.time()

  return(list(
    est_df = bind_rows(psm_list),
    fail_count = fail_count,
    comp_time = as.numeric(difftime(end_time, start_time, units = 'secs'))/(n-fail_count))
  )
}

psm_sim_out <- psm_sim(sim_list) 
psm_sim_linear <- psm_sim(sim_linear)

psm_df <- psm_sim_out$est_df %>%
  mutate(fail_count = psm_sim_out$fail_count,
         comp_time = psm_sim_out$comp_time)

psm_df_linear <- psm_sim_linear$est_df %>%
  mutate(fail_count = psm_sim_linear$fail_count,
         comp_time = psm_sim_linear$comp_time)

write_csv(psm_df, here('files', 'psm_df.csv'))
write_csv(psm_df_linear, here('files', 'psm_df_linear.csv'))
```

```{r aipw-sim, eval = F}
aipw_sim <- function(dat, seed = 185){

  start_time <- Sys.time()

  set.seed(seed)
  n <- length(dat)
  aipw_list <- list()
  # fail_count <- 0

  for(i in 1:n){
    print(paste0(i, ' out of ', n))
    sim_dat <- data.frame(y = dat[[i]]$y, 
                          d = as.numeric(ifelse(dat[[i]]$z == 'trt', 1, 0)), 
                          dat[[i]]$x)

    tryCatch({
      sim_dat <- sim_dat %>%
        fastDummies::dummy_cols(remove_first_dummy = T, remove_selected_columns = T) 
      }, error=function(e){
      })

    tryCatch({
      forest <- grf::causal_forest(X = select(sim_dat, 3:length(names(sim_dat))), 
                                   Y = sim_dat$y, W = sim_dat$d)
      # forest <- grf::causal_forest(X = select(sim_dat, starts_with('x')), 
      #                              Y = sim_dat$y, W = sim_dat$d)

      aipw_out <- grf::average_treatment_effect(forest, target.sample = 'treated', method = 'AIPW')

      aipw_list[[i]] <- data.frame(d = aipw_out[[1]],
                                   se = aipw_out[[2]],
                               truth = mean(dat[[i]]$y.1) - mean(dat[[i]]$y.0))

  }, error=function(e){
    cat("ERROR :",conditionMessage(e), "\n")
    })
  }

  fail_count <- sum(sapply(aipw_list, function(x) is.null(x)))
  # in case last few are errors
  fail_count <- ifelse(length(aipw_list) == n, fail_count, fail_count + (n - length(aipw_list)))

  end_time <- Sys.time()

  return(list(
      est_df = bind_rows(aipw_list),
      fail_count = fail_count,
      comp_time = as.numeric(difftime(end_time, start_time, units = 'secs'))/(n-fail_count))
  )
}

aipw_out <- aipw_sim(sim_list)
aipw_out_linear <- aipw_sim(sim_linear)

aipw_df <- aipw_out$est_df %>%
  mutate(fail_count = aipw_out$fail_count,
         comp_time = aipw_out$comp_time)

aipw_df_linear <- aipw_out_linear$est_df %>%
  mutate(fail_count = aipw_out_linear$fail_count,
         comp_time = aipw_out_linear$comp_time)

write_csv(aipw_df,  here('files', 'aipw_df.csv'))
write_csv(aipw_df_linear,  here('files', 'aipw_df_linear.csv'))

@@

tmle_sim <- function(dat, seed = 185){
  start_time <- Sys.time()

  set.seed(seed)
  n <- length(dat)
  tmle_list <- list()
  # fail_count <- 0

  for(i in 1:n){
    print(paste0(i, ' out of ', n))
    sim_dat <- data.frame(y = dat[[i]]$y, 
                          d = ifelse(dat[[i]]$z == 'trt', 1, 0), 
                          dat[[i]]$x)

    tryCatch({
      sim_dat <- sim_dat %>%
        fastDummies::dummy_cols(remove_first_dummy = T, remove_selected_columns = T) 
      }, error=function(e){
      })

    tryCatch({
      forest <- grf::causal_forest(X = select(sim_dat, 3:length(names(sim_dat))), 
                                   Y = sim_dat$y, W = sim_dat$d)
      # forest <- grf::causal_forest(X = select(sim_dat, starts_with('x')), 
      #   Y = sim_dat$y, W = sim_dat$d)

      tmle_out <- grf::average_treatment_effect(forest, target.sample = 'treated', method = 'TMLE')

      tmle_list[[i]] <- data.frame(d = tmle_out[[1]],
                                   se = tmle_out[[2]],
                               truth = mean(dat[[i]]$y.1) - mean(dat[[i]]$y.0))

  }, error=function(e){
    cat("ERROR :",conditionMessage(e), "\n")
    })
  }

  fail_count <- sum(sapply(tmle_list, function(x) is.null(x)))
  # in case last few are errors
  fail_count <- ifelse(length(tmle_list) == n, fail_count, fail_count + (n - length(tmle_list)))

  end_time <- Sys.time()

 return(list(
      est_df = bind_rows(tmle_list),
      fail_count = fail_count,
      comp_time = as.numeric(difftime(end_time, start_time, units = 'secs'))/(n-fail_count))
  )
}

tmle_out <- tmle_sim(sim_list)
tmle_out_linear <- tmle_sim(sim_linear)

tmle_df <- tmle_out$est_df %>%
  mutate(fail_count = tmle_out$fail_count,
         comp_time = tmle_out$comp_time)

tmle_df_linear <- tmle_out_linear$est_df %>%
  mutate(fail_count = tmle_out_linear$fail_count,
         comp_time = tmle_out_linear$comp_time)

write_csv(tmle_df, here('files', 'tmle_df.csv'))
write_csv(tmle_df_linear, here('files', 'tmle_df_linear.csv'))
```

@@

  start_time <- Sys.time()

@@

  dml_list <- list()

  for(i in 1:n){
    print(paste0(i, ' out of ', n))
    sim_dat <- data.frame(y = dat[[i]]$y, 
                          d = ifelse(dat[[i]]$z == 'trt', 1, 0), 
                          dat[[i]]$x)

@@

    tryCatch({
      dml_out <- DoubleML::DoubleMLPLR$new(
        DoubleML::DoubleMLData$new(sim_dat,
                                 y_col = 'y',
                                 d_cols = 'd',
                                 x_cols = names(dat[[i]]$x)), 
        ml_l=ml_l, ml_m=ml_m)

      dml_out$fit()
      dml_list[[i]] <- data.frame(d = dml_out$all_coef[[1,1]],
                                  se = dml_out$all_se[[1,1]],
                               truth = mean(dat[[i]]$y.1) - mean(dat[[i]]$y.0))

  }, error=function(e){
    cat("ERROR :",conditionMessage(e), "\n")
    })
  }

  fail_count <- sum(sapply(dml_list, function(x) is.null(x)))
  # in case last few are errors
  fail_count <- ifelse(length(dml_list) == n, fail_count, fail_count + (n - length(dml_list)))

  end_time <- Sys.time()

  return(list(
      est_df = bind_rows(dml_list),
      fail_count = fail_count,
      comp_time = as.numeric(difftime(end_time, start_time, units = 'secs'))/(n-fail_count))
      )
}

dml_out <- dml_sim(sim_list)
dml_out_linear <- dml_sim(sim_linear)

dml_df <- dml_out$est_df %>%
  mutate(fail_count = dml_out$fail_count,
         comp_time = dml_out$comp_time)
dml_df_linear <- dml_out_linear$est_df %>%
  mutate(fail_count = dml_out_linear$fail_count,
         comp_time = dml_out_linear$comp_time)

write_csv(dml_df, here('files', 'dml_df.csv'))
write_csv(dml_df_linear, here('files', 'dml_df_linear.csv'))
```

```{r dorie-results}
bind_rows(
  perform(read_csv(here('files', 'lm_df.csv')), 'lm'),
  perform(read_csv(here('files', 'psm_df.csv')), 'psm'),
  perform(read_csv(here('files', 'aipw_df.csv')), 'aipw'),
  perform(read_csv(here('files', 'tmle_df.csv')), 'tmle'),
  perform(read_csv(here('files', 'dml_df.csv')), 'dml')
  ) %>%
  kableExtra::kable(booktabs = T, 
                    digits = 3,
                    linesep = '',
                    caption = 'Results of Monte Carlo simulations using the first 20 datasets from Dorie et al. (2019), 5 replications each. Percent bias is calculated as the estimator\'s bias as a percentage of its standard error, rmse is root mean squared error, mae is median absolute error, and comp\\_time is average computation time measured in seconds for each dataset.')
```

```{r dorie-results-linear}
bind_rows(
  perform(read_csv(here('files', 'lm_df_linear.csv')), 'lm'),
  perform(read_csv(here('files', 'psm_df_linear.csv')), 'psm'),
  perform(read_csv(here('files', 'aipw_df_linear.csv')), 'aipw'),
  perform(read_csv(here('files', 'tmle_df_linear.csv')), 'tmle'),
  perform(read_csv(here('files', 'dml_df_linear.csv')), 'dml')
  ) %>%
  kableExtra::kable(booktabs = T, 
                    digits = 3,
                    linesep = '',
                    caption = 'Results of Monte Carlo simulations using the linear DGP for dataset 3 from Dorie et al. (2019), 5 replications each. Percent bias is calculated as the estimator\'s bias as a percentage of its standard error, rmse is root mean squared error, mae is median absolute error, and comp\\_time is average computation time measured in seconds for each dataset.')
```

Results are shown in Table \@ref(tab:dorie-results). The methods all have similar levels of bias and error, but AIPW performs better than the other methods on all measures. AIPW did fail, however, on the most data sets; the package `CausalGAM` package does not do well with lack of overlap of factor variables between treated and control groups.

```{r lalonde-data}
unemp_func <- function(x){
  x %>%
    as.data.frame() %>%
    mutate(re74_0 = re74 == 0,
         re75_0 = re75 == 0) 
}

lalonde_exp <- read_dta(here('data', 'nsw.dta')) %>%
  as.data.frame() %>%
  mutate(re75_0 = re75 == 0)
lalonde_exp_74 <- read_dta(here('data', 'nsw_dw.dta')) %>%
  unemp_func()
lalonde_cps1_controls <- read_dta(here('data', 'cps_controls.dta')) %>%
  unemp_func()
lalonde_cps3_controls <- read_dta(here('data', 'cps_controls3.dta')) %>%
  unemp_func()
lalonde_psid1_controls <- read_dta(here('data', 'psid_controls.dta')) %>%
  unemp_func()
lalonde_psid3_controls <- read_dta(here('data', 'psid_controls3.dta')) %>%
  unemp_func()
```

## Comparisons using LaLonde NSW Data

As another benchmark test, I use data from LaLonde's [-@lalonde_1986_evaluating] study of the National Supported Work Demonstration (NWS), as provided by @dehejia_1999_causal. The NWS randomly provided training to disadvantaged workers, allowing an experimental estimate of the effect of the intervention. This study had `r nrow(filter(lalonde_exp, treat == T))` treated and `r nrow(filter(lalonde_exp, treat == F))` control participants
